### 2010-03-23 papers/icfp-2010-desc/paper_idesc.tex
Desc@ICFP: evalOne: fix typo
Desc@ICFP: vectors have finally landed.

While inductive types are the alpha and omega of simply-typed
languages, evolving in a dependent setting fosters new
opportunities. The typical example is bounded lists, also known as
vectors. A vector is a list decorated by its length. Having this
information prevents hazardous operations, such as taking the head of
an empty vector: the head function only takes vectors of length
$\NatSuc{n}$, as enforced by its type. This is made possible by the
very specificity of dependently-typed systems: terms -- the length --
are allowed to influence types -- the vector type.

@@

In the case of vectors, for instance, for the index $\NatZero$, we
would only propose the constructor $\ListNil$. Similarly, for
$\NatSuc{n}$, we only propose the constructor $\ListCons{}{}$. We will
see more instances of this pattern in the following.
\note{We're talking about vectors again, are we \emph{really} not going to
define them?}

@@

\[\stk{
\evalOne : \PI{\V{ty}}{\Ty} \idescop{\toIDesc{\HExprD}\: ty}{\Ty}{ty} \To {\Val{ty}} \\
\begin{array}{@{}l@{\:=\:\:}l}
\evalOne\: \_\: \pair{\Eval{}}{x}{}                                         & x \\
\evalOne\: \_\: \pair{\Econd{}{}{}}{\pair{\BoolTrue}{\pair{x}{\_}{}}{}}{}   & x \\
\evalOne\: \_\: \pair{\Econd{}{}{}}{\pair{\BoolFalse}{\pair{\_}{y}{}}{}}{}  & x \\
\evalOne\: \Enat\: \pair{\Eplus{}{}}{\pair{x}{y}{}}{}                       & \F{plusHost}\: x y \\
\evalOne\: \Ebool\: \pair{\Ele{}{}}{\pair{x}{y}{}}{}                        & \F{leHost}\: x y 
\end{array} \\
\\
\evalH : \PI{\V{ty}}{\Ty} 
           \IMu{\Ty}{\toIDesc{\HExprD}}{ty} \To
           \Val{ty} \\
\evalH\: ty\: term = \F{cataI}\: \Ty\: 
                                 \toIDesc{\HExprD}\: 
                                 \Val{}\: 
                                 \evalOne\: 
                                 ty\: 
                                 term
}\]
\note{pwm: $\evalOne$ rather illustrates my point about the 2 types of
  IFunc floating around, $\idescop{\_}{}{}$ is defined using the 1st
  notion, and used it here as if it were the 2nd. pierre: in this
  case, it was a typo. But you're right, there is two notions and we
  probably want to clarify where these things live.}

@@

\note{ I suspect we don't implement vectors, space-wise }
\note{pwm: I suspect we don't get away with leaving them out}

>>>>

While inductive types are the alpha and omega of simply-typed
languages, evolving in a dependent setting fosters new
opportunities. The typical example is bounded lists, also known as
vectors. A vector is a list decorated by its length. Having this
information prevents hazardous operations, such as taking the head of
an empty vector: the head function only takes vectors of length
$\NatSuc{n}$, as enforced by its type. This is made possible by the
very specificity of dependently-typed systems: terms -- the length --
are allowed to influence types -- the vector type. \note{Reference
  paper to cite about vectors in dependent types?}

@@

In the case of vectors, for instance, for the index $\NatZero$, we
would only propose the constructor $\ListNil$. Similarly, for
$\NatSuc{n}$, we only propose the constructor $\ListCons{}{}$. Hence,
we obtain the following definition:

\newcommand{\VecD}{\C{VecD}}
\newcommand{\VecNil}{\etag{\CN{vnil}}}
\newcommand{\VecCons}[2]{\etag{\CN{vcons}}\:#1\:#2}

\[\stk{
\VecD : \Set \To \TagIDesc{\Nat} \\
\begin{array}{@{}ll}
\VecD\: X = \bigRedBracket{ 
            \begin{array}{l}
                \pair{[]}{\_}{} \\
                \bigRedBracket{
                \begin{array}{l} 
                  \LAM{n} \caseB{n}{\NatZero \To [\VecNil] \\
                                   (\NatSuc n) \To [\VecCons{}{}]} \\
                  \LAM{n} \caseB{n}{\NatZero \To [\DConst{Unit}] \\
                                   (\NatSuc n) \To [\DSigma{X}{\DVar{n}}]}
                \end{array}}
            \end{array}}
\end{array}
}\]

We will see more instances of this pattern in the following.

@@

\[\stk{
\evalOne : \PI{\V{ty}}{\Ty} \idescop{\toIDesc{\HExprD}\: ty}{\Ty}{ty} \To {\Val{ty}} \\
\begin{array}{@{}l@{\:=\:\:}l}
\evalOne\: \_\: \pair{\Eval{}}{x}{}                                         & x \\
\evalOne\: \_\: \pair{\Econd{}{}{}}{\pair{\BoolTrue}{\pair{x}{\_}{}}{}}{}   & x \\
\evalOne\: \_\: \pair{\Econd{}{}{}}{\pair{\BoolFalse}{\pair{\_}{y}{}}{}}{}  & x \\
\evalOne\: \Enat\: \pair{\Eplus{}{}}{\pair{x}{y}{}}{}                       & \F{plusHost}\: x\: y \\
\evalOne\: \Ebool\: \pair{\Ele{}{}}{\pair{x}{y}{}}{}                        & \F{leHost}\: x\: y 
\end{array} \\
\\
\evalH : \PI{\V{ty}}{\Ty} 
           \IMu{\Ty}{\toIDesc{\HExprD}}{ty} \To
           \Val{ty} \\
\evalH\: ty\: term = \F{cataI}\: \Ty\: 
                                 \toIDesc{\HExprD}\: 
                                 \Val{}\: 
                                 \evalOne\: 
                                 ty\: 
                                 term
}\]
\note{pwm: $\evalOne$ rather illustrates my point about the 2 types of
  IFunc floating around, $\idescop{\_}{}{}$ is defined using the 1st
  notion, and used it here as if it were the 2nd. pierre: in this
  case, it was a typo. But you're right, there is two notions and we
  probably want to clarify where these things live.}

@@
